---
title: "p8105_hw3_jy3186"
author: "Jiayi Yang"
date: "2022-10-11"
output: github_document
---

```{r}
library(tidyverse)
library(dplyr)
library(rnoaa)
library(ggridges)
library(patchwork)
library(p8105.datasets)

```

##Problem 1 from Jeff's solution
```{r setup, include=FALSE}


knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Due date

Due: October 15 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |
| Optional survey | No points |


### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```





### Problem 2

load and tidy the data 
```{r}
accel_data = read_csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    weekday_weekend = ifelse(day %in% c("Saturday","Sunday"), "Weekend", "Weekday"),
    day_id = as.integer(day_id),
    week = as.integer(week)
  ) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    values_to = "activity_value"
  ) %>% 
  relocate(weekday_weekend)
accel_data
```
There are `r nrow(accel_data)` observations and `r ncol(accel_data)` variables.
There are four date variables `weekday_weekend`, `week`, `day_id` and `day`; and there are `activity_` indicates the activity minutes;
and there are 1440 variables of activity observations.

aggregate activities and make a plot to show the trend
```{r}
accel_data %>% 
  group_by(day_id) %>% 
  mutate(
    activity_total =  sum(activity_value, na.rm = TRUE),
    activity_total = as.integer(activity_total)
  ) %>% 
  pivot_wider(
    names_from = "day",
    values_from = "activity_total"
  ) %>% 
  select(Monday, Tuesday, Wednesday, Thursday, Friday, Saturday, Sunday) %>% 
  knitr::kable()
```
There are no specific trend of the data shown from the table here.

Make a single-panel plot that shows the 24-hour activity time courses for each day
```{r}
accel_eachday = accel_data %>% 
  ggplot(aes(x = activity, y = activity_value, color = day), se = FALSE) + 
  geom_line(alpha = .5) + geom_smooth(aes(group = day), se = FALSE) +
    labs(
    title = "24-hour activity plot for each day",
    x = "minute",
    y = "day"
  ) + 
    theme(legend.position = "bottom")
accel_eachday
```
This patient has a pattern to have higher activity minutes during weekdays than weekends.


### Problem 3

load data from web
```{r}
library(p8105.datasets)
data("ny_noaa")

```

The data contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns.
The data has weather information from 1981/1/1 to 2010/12/31, with variables such as prcp, snow, snwd, tmax, and tmin.

The units for each variable are:
`id` weather station id;
`prcp` precipitation (tenths of mm);
`snwd` snow depth (mm);
`tmax` maximum temperature (tenths of Celsius degrees);
`tmin` minimum temperature (tenths of Celsius degrees); 


```{r}
ny_noaa_df = 
  ny_noaa %>%
 mutate(
    tmax = as.integer(tmax),
    tmin = as.integer(tmin),
   tmax = tmax/10,
   tmin = tmin/10,
   prcp = prcp/10
  ) %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(
    month = as.integer(month),
    day = as.integer(day)
  )
ny_noaa_df

```


count snowfall
```{r}

ny_noaa_df %>% 
  count(snow, na.rm = TRUE) %>% 
  arrange(desc(n))
```
The most observed snowfall value is 0 with a 2008508 frequency of happening. Because there is usually no snow in most days of a year.

#### January and July average max temperature
```{r, message= FALSE}
Jan_July_df = ny_noaa_df %>% 
  filter(month == 1 | month == 7) %>% 
  group_by(id, month, year) %>% 
  mutate(tmax_mean = mean(tmax, na.rm = TRUE),
         tmax_mean = as.integer(tmax_mean)) %>% 
  summarize(tmax_mean)

Jan_July_df
ggplot(Jan_July_df, aes(x = year, y = tmax_mean), color = month) +
 geom_point() + geom_line(alpha = .5) + geom_smooth(aes(group = month), se = FALSE) +
  labs(
    title = "Average Max temperature of Jan and July from 1980 to 2010",
    x = "Year",
    y = "Average max temperature (C)"
  ) + 
    theme(legend.position = "bottom") +
  facet_grid(. ~month)

```
There are several outliers in both January and July average max temperatures. The observable structure for Janurary is wavy and for July is more toward a line (more linear).

#### Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); 
```{r}
tmax_vs_tmin = 
  ny_noaa_df %>% 
  drop_na(tmax, tmin) %>% 
  select(id, year, month, day, tmax, tmin) %>% 
  ggplot(aes(x = tmin, y = tmax)) + geom_hex() +
  labs(
    title = "Temperature max vs min from 1980 to 2010",
    x = "tmin",
    y = "tmax"
  ) + 
    theme(legend.position = "bottom")

tmax_vs_tmin
```
## Make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year
```{r}
snowfall_plot =
  ny_noaa_df %>% 
  select(id, year, month, day, snow) %>%
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x= snow, fill = year)) +
  geom_density(color = "blue", alpha = .5) +
    labs(
    title = "Snowfall distribution (0-100) from 1980 to 2010",
    x = "snowfall",
    y = "snowfall density"
  ) + 
    theme(legend.position = "bottom")


snowfall_plot
```
```{r}
tmax_vs_tmin + snowfall_plot
```



